### 2024-10-18 archive.qmd

---
title: "text dump"
format: html
---

<!-- Fix cluster-level missingness determinately. -->
```{r}
# meth <- make.method(dat)
# meth["experience_j"] <- "2lonly.mean"
# meth[meth != "2lonly.mean"] <- ""
# pred <- make.predictorMatrix(dat)
# pred[] <- 0
# pred[, "cluster_id"] <- -2
# imp <- mice(
#   popular_MAR, 
#   m = 1, 
#   method = meth, 
#   predictorMatrix = pred,
#   maxit = 1)
# plot_pattern(complete(imp))
```

This procedure has fixed the missing experience value

Does not work if there are inconsistencies within clusters! But we can do this manually.

```{r}
# dat |>
#   group_by(cluster_id) |>
#   reframe(teacher_experience = unique(experience_j))
```

This shows us that in the first cluster teacher experience of 4 and 24 is recorded.

```{r}
# dat |>
#   filter(cluster_id == 1) |>
#   ggplot(aes(experience_j)) + geom_histogram()
```

For most pupils, teacher experience is 24. We can quite safely assume that the teacher experience of 4 is a data entry error. We can fix this by over-imputing the teacher experience of 4 to 24, and deductively imputing the missing teacher experience from the same cluster as 24. The following line of code sets the teacher experience of 4 manually as missing.

```{r}
# dat[dat$cluster_id == 1 & (is.na(dat$experience_j) | dat$experience_j == 4), "experience_j"] <- NA
```

```{r}
# imp <- mice(
#   popular_MAR, 
#   m = 1, 
#   method = meth, 
#   predictorMatrix = pred,
#   maxit = 1)
# plot_pattern(complete(imp))
```